---
title: 'Milestone #5'
author: "Dominic Skinnion"
date: "10/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr)
library(janitor)
library(plotrix)
library(scales)
library(ggthemes)
library(robustfa)
library(tidyverse)
```
# About

In the United States, the people do not directly vote for President. The Founding Fathers believed that a true democracy was subject to the rule of factions and mob rule. The electoral college, they believed, would protect against this. Instead of having each person's vote count towards President, each state receives a certain number of electoral votes. The number of electoral votes for each state is equal to the numbers of members of congress for that state (the number of senators, plus the number of representatives). When the 23rd Amendment was ratified in 1961, Washington D.C. was also given electoral votes, but not to exceed the lowest number of votes that a state has -- this has usually limited D.C. to 3 votes.

The data for the United States general elections was found from a Github repository which included multiple government datasets. Elections that were contested were labelled with asterisks in this dataset, so I ammended the dataset with the wikipedia data set found here: https://en.wikipedia.org/wiki/United_States_Electoral_College. The population data for the states are based off of the United States Census, conducted every 10 years. This dataset was found from a GitHub repository with historical data.

```{r data, include = FALSE}

elections <- read.csv("raw-data/elec.csv") %>%
   separate("Year", c("year", "president"), sep = "-", remove = TRUE,
  convert = TRUE) %>%
  clean_names()


census <- read.csv("raw-data/nhgis0011_ts_state.csv") %>%
  clean_names()
```

```{r elections data, include = FALSE}
# colnames(elections)[colnames(elections)=="x_3"] <- "percent_4"
# colnames(elections)[colnames(elections)=="ev_3"] <- "ev_4"
# colnames(elections)[colnames(elections)=="x_2"] <- "percent_3"
# colnames(elections)[colnames(elections)=="ev_2"] <- "ev_3"
# colnames(elections)[colnames(elections)=="x_1"] <- "percent_2"
# colnames(elections)[colnames(elections)=="ev_1"] <- "ev_2"
# colnames(elections)[colnames(elections)=="x"] <- "percent_1"
# colnames(elections)[colnames(elections)=="ev"] <- "ev_1"
```

```{r fixing variable types, echo = FALSE, warning = FALSE}
elections <- elections %>%
  mutate(year = year, president = president, state = state, total = total, party_1 = as.numeric(as.character(party_1)), percent_1 = as.numeric(as.character(x)), electoral_votes_1 = as.numeric(as.character(ev)), party_2 = as.numeric(as.character(party_2)), percent_2 = as.numeric(as.character(x_1)), electoral_votes_2 = as.numeric(as.character(ev_1)), party_3 = as.numeric(as.character(party_3)), percent_3 = as.numeric(as.character(x_2)), electoral_votes_3 = as.numeric(as.character(ev_2)), party_4 = as.numeric(as.character(party_4)), percent_4 = as.numeric(as.character(x_3)), electoral_votes_4 = as.numeric(as.character(ev_3)), party = party, notes = notes) %>%
  select(year, president, state, total, party_1, percent_1, electoral_votes_1, party_2, percent_2, electoral_votes_2, party_3, percent_3, electoral_votes_3, party_4, percent_4, electoral_votes_4, party, notes)

elections[is.na(elections)] <- 0
```

```{r adding total EV, echo = FALSE}
elections_with_total_ev <- elections %>%
  mutate(electoral_votes_total = electoral_votes_1 + electoral_votes_2 + electoral_votes_3 + electoral_votes_4) %>%
  mutate(electoral_votes_total = ifelse(electoral_votes_total == 0, party_1 + party_2 + party_3 + party_4, electoral_votes_total))
```

```{r fixing total votes, echo = FALSE}
electoral_votes <- elections_with_total_ev %>%
  select(year, president, state, party, electoral_votes_total)

# Need to fix numbers: contested elections. I am looking at the number of electoral votes allotted, so these numbers should still be counted. Corrected numbers from https://en.wikipedia.org/wiki/United_States_Electoral_College
electoral_votes[488, 5] = 6
electoral_votes[493, 5] = 11
electoral_votes[498, 5] = 12
electoral_votes[499, 5] = 8
electoral_votes[501, 5] = 8
electoral_votes[506, 5] = 15
electoral_votes[518, 5] = 12
electoral_votes[519, 5] = 8
electoral_votes[1429, 5] = 11
electoral_votes[1477, 5] = 11
electoral_votes[1500, 5] = 8
electoral_votes[1512, 5] = 8
electoral_votes[1675, 5] = 12
electoral_votes[1727, 5] = 9
electoral_votes[1881, 5] = 6
electoral_votes[1994, 5] = 3
electoral_votes[2060, 5] = 10
electoral_votes[2115, 5] = 5
electoral_votes[2201, 5] = 4
electoral_votes[2233, 5] = 38
electoral_votes[2237, 5] = 12
```

```{r state, include = FALSE}
# 
# separate_by_state <- function(n){
# 
#    nf  <- gsub("()","",sys.call()[51])
#    titles <- c(Alambama = "Voters in Alabama",
#                Alaska = "Voters in Alaska",
#                Arizona = "Voters in Arizona",
#                Arkansas = "Voters in Arkansas",
#                California = "Voters in California",
#                Colorado = "Voters in Colorado",
#                Connecticut = "Voters in Connecticut",
#                Delaware = "Voters in Delaware",
#                Dist. of Col. = "Voters in District of Columbia",
#                Florida = "Voters in Florida",
#                Georgia = "Voters in Georgia",
#                Hawaii = "Voters in Hawaii",
#                Idaho = "Voters in Idaho",
#                Illinois = "Voters in Illinois",
#                Indiana = "Voters in OIndiana",
#                Iowa = "Voters in Iowa",
#                Kansas = "Voters in Kansas",
#                Kentucky = "Voters in Kentucky",
#                Louisiana = "Voters in Louisiana",
#                Maine = "Voters in Maine",
#                Maryland = "Voters in Maryland",
#                Massachusetts = "Voters in Massachusetts",
#                Michigan = "Voters in Michigan",
#                Minnesota = "Voters in Minnesota",
#                Mississippi = "Voters in Mississippi",
#                Missouri = "Voters in Missouri",
#                Montana = "Voters in Montana",
#                Nebraska = "Voters in Nebraska",
#                Nevada = "Voters in Nevada",
#                New Hampshire = "Voters in New Hampshire",
#                New Jersey = "Voters in New Jersey",
#                New Mexico = "Voters in New Mexico",
#                New York = "Voters in North Carolina",
#                North Carolina = "Voters in North Carolina",
#                North Dakota = "Voters in North Dakota",
#                Ohio = "Voters in Ohio",
#                Oklahoma = "Voters in Oklahoma",
#                Oregon = "Voters in Oregon",
#                Pennsylvania = "Voters in Pennsylvania",
#                Rhode Island = "Voters in Rhode Island",
#                South Carolina = "Voters in South Carolina",
#                South Dakota = "Voters in South Dakota",
#                Tennessee = "Voters in Tennessee",
#                Texas = "Voters in Texas",
#                Utah = "Voters in Utah",
#                Vermont = "Voters in Vermont",
#                Virginia = "Voters in Virginia",
#                Washington = "Voters in Washington",
#                West Virginia = "Voters in West Virginia",
#                Wisconsin = "Voters in Wisconsin",
#                Wyoming = "Voters in Wyoming"
#                )
# 
#    plot <- elections %>%
#      filter(state == n) %>%
#      select(year, total) %>%
#      ggplot(aes(x = year, y = total)) +
#        labs(x = "Year", y = "Number of Voters") +
#        geom_line() +
#        ggtitle(titles[nf])
# 
#     return(plot)
# }
# 
# states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Dist. of Col.", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming")
# 
# map(states, separate_by_state)
```
# 10 States with the Most Electoral Votes in 2016

```{r ev 2016 plot, echo = FALSE}
electoral_votes %>%
  filter(year == 2016) %>%
  arrange(desc(electoral_votes_total)) %>%
  slice(1:10) %>%
  ggplot(aes(x = reorder(state, -electoral_votes_total), y = electoral_votes_total, fill = party)) +
    geom_col() +
    coord_flip() +
    scale_fill_manual(values = c("#0015BC", "#BC1500"), name = "Party") +
    labs(title = "States with the Most Electoral Votes in 2016", x = "Number of Electoral Votes", y = "State")
```


# Number of Voters Over Time in the 13 Original Colonies
```{r MA, echo = FALSE}

MA <- elections %>%
  filter(state == "Massachusetts")

ggplot(MA, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in Massachusetts", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r DE, echo = FALSE}
DE <- elections %>%
  filter(state == "Delaware")

ggplot(DE, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in Delaware", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r MD, echo = FALSE}
MD <- elections %>%
  filter(state == "Maryland")

ggplot(MD, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in Maryland", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r CT, echo = FALSE}
CT <- elections %>%
  filter(state == "Connecticut") 

ggplot(CT, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in Connecticut", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r RI, echo = FALSE}
RI <- elections %>%
  filter(state == "Rhode Island")

ggplot(RI, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in Rhode Island", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r NH, echo = FALSE}
NH <- elections %>%
  filter(state == "New Hampshire") 

ggplot(NH, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in New Hampshire", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r NY, echo = FALSE}
NY <- elections %>%
  filter(state == "New York") 

ggplot(NY, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in New York", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r PA, echo = FALSE}
PA <- elections %>%
  filter(state == "Pennsylvania")

ggplot(PA, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Pennsylvania", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r NJ, echo = FALSE}
NJ <- elections %>%
  filter(state == "New Jersey") 

ggplot(NJ, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in New Jersey", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r VA, echo = FALSE}
VA <- elections %>%
  filter(state == "Virginia")

ggplot(VA, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in Virginia", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r NC, echo = FALSE}
NC <- elections %>%
  filter(state == "North Carolina") 

ggplot(NC, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in North Carolina", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r SC, echo = FALSE}
SC <- elections %>%
  filter(state == "South Carolina") 

ggplot(SC, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in South Carolina", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```

```{r GA, echo = FALSE}
GA <- elections %>%
  filter(state == "Georgia") 

ggplot(GA, aes(x = year, y = total)) +
  geom_line() +
  labs(title = "Voters in Georgia", x = "Year", y = "Number of Voters") +
  scale_y_continuous(labels = comma, limits = c(0, 8000000)) +
  theme_stata()
```
